---
title: "NS Hemisphere Analysis"
author: "KE Lotterhos"
date: "10/27/2020"
output: html_document
---


Based on basic code for calculating and mapping climatic novelty using the sigma dissimilarity metric. 
* written by Colin Mahony
* Supp Material of "A closer look at novel climates: new methods and insights at continental to landscape scales"

# What this code does
### Compare 1800 analog (A) to today surface (B), with today's surface variability (C)
### Compare today surface analog (A) to 2100-RCP8.5 (B), with today's surface variability (C)
### Compare today surface analog (A) to 2100-RCP4.5 (B), with today's surface variability (C)

## This code also compares the converse (e.g. the current climate as B to the future climate as A) to assess disappearing climates

This code uses the annnual climate normals and monthly ICV for the novelty calculation

```{r}
##################################
#### System setup ####
##################################
## Specify location of data ####
setwd("XXXX") 
#setwd("~/Desktop/PostDoc/OceanClimateNovelty/")

source("src/0-Novelty_Oceans_Functions.R")
```

```{r}
##################################
#### Read in the input data ####
##################################

# set path to data
path <- "../Data/model/"

dat1 <- fread(paste0(path,"Temp_Arag_1800_2000.txt"), sep = ",")
dat4.5 <- fread(paste0(path,"Temp_Arag_2070_2100_RCP45.txt"), sep = ",")
dat8.5 <- fread(paste0(path,"Temp_Arag_2070_2100_RCP85.txt"), sep = ",")

ICV <- fread(paste0(path, "ICV_Temp_Arag_pH_1960_2020.txt"))

results_path <- "../results/" # you may have to create this

results_dir <- paste0(results_path, "HemisphereRestricted/")

head(dat1)
head(dat8.5)
head(dat4.5)
head(ICV)
unique(dat1$Year)
unique(dat8.5$Year)
unique(dat4.5$Year)
unique(ICV$Year)

hist(c(dat1$Arag, dat8.5$Arag))
hist(log10(c(dat1$Arag, dat8.5$Arag)))
```

## boxplot ICV
```{r, boxplot ICV}
pdf(paste0(results_dir, "ICV_Lat.pdf"), width=9, height=7)
par(mfrow=c(3,1), mar=c(4,5,0.5,0.5), oma=c(2,0,0,0), las=0)

lat <- sort(unique(ICV$Lat))
length(lat)
lat[1]
lat[42]
lat[80]
lat[117]
lat[158]
ind <- c(1,42,80,117,158)
lat_names=rep(NA, 169)
lat_names[ind] <- lat[ind]
lat_names[80] <- 0 #close enough

boxplot(ICV$SST~ICV$Lat, col="cyan",  
        #border = rgb(0,0,0,0.3),  outpch=20, outcol= rgb(0,0,0,0.3), outbg=rgb(0,0,0,0.3), 
        ylab="", xlab="", names=lat_names, las=1, cex=1.2, outline=FALSE, border=rgb(0,0,0,0.5), lty=1)
  mtext("SST", side=2, cex=1.5, line=3)

boxplot(ICV$Arag~ICV$Lat, col="coral",
        ylab="", xlab="", names=lat_names, las=1, cex=1.2, outline=FALSE, border=rgb(0,0,0,0.5), lty=1)
  mtext("Arag.", side=2, cex=1.5, line=3)

boxplot(ICV$pH ~ ICV$Lat, col="gold",
        ylab="", xlab="", names=lat_names, las=1, cex=1.2, outline=FALSE, border=rgb(0,0,0,0.5), lty=1)
  mtext("pH", side=2, cex=1.5, line=3)


mtext("Latitude", side=1, outer=TRUE, cex=2)
dev.off()
```


```{r}
##################################
#### Log-transform Arag ####
##################################
dat1$Arag <- log10(dat1$Arag)
dat4.5$Arag <- log10(dat4.5$Arag)
dat8.5$Arag <- log10(dat8.5$Arag)
ICV$Arag <- log10(ICV$Arag)

#  Aragonite saturation is a ratio variable; it is limited at zero and 
# proportional changes are meaningful. The analysis uses raw values of 
# aragonite saturation state. Under raw scaling, the difference between 1 and 2 
# is equal to the difference between zero and one, which doesn’t reflect the 
# meaning of the variable. Instead, the difference between 1 and 2 should 
# be given the same significance as the difference between 0.5 and 1 (doubling vs halving). 
# Log-transforming aragonite saturation state will produce this more meaningful 
# proportional scaling.

# More generally on the same topic: The need for proportional scaling is the 
# reason why precipitation and other ratio variables (e.g. degree-days) are 
# typically log-transformed in climate space analysis. In theory, all variables 
# should be log-transformed to produce an environmental space where distances 
# are comparable for different locations; however, in practice temperature 
# doesn’t need to be log-transformed because it doesn’t vary across orders 
# of magnitude, and in our case pH is already a log-scaled variable.

```



```{r}
#--------------------------------
#### 40-year climate normals ####
#--------------------------------
# for example, 1930 represents from 1/1/1925 to 12/31/1934.  
dat_2000 <-   dat1 %>% filter(Year>1960 & Year<2010)
dim(dat_2000)

dat_1800 <- dat1 %>% filter(Year<1850)
dim(dat_1800)

dat_2100_8.5 <-   dat8.5 
dat_2100_4.5 <-   dat4.5 

sum(!complete.cases(dat_1800))
sum(!complete.cases(dat_2000))
sum(!complete.cases(dat_2100_4.5))
sum(!complete.cases(dat_2100_8.5))
sum(!complete.cases(ICV))

# Calculate climate normals
# Returns the mean climate for each station at each time
norm_1800 <- calculate_normals_annual(dat_1800)
norm_2000 <- calculate_normals_annual(dat_2000)
norm_2100_8.5 <- calculate_normals_annual(dat_2100_8.5)
norm_2100_4.5 <- calculate_normals_annual(dat_2100_4.5)

head(norm_1800)
dim(norm_1800)
dim(norm_2000)
dim(norm_2100_8.5)
dim(norm_2100_4.5)
```

```{r}
#--------------------------------  
### data frame to link station number to Lat Long ####
#--------------------------------
head(dat8.5)
stations <- unique(dat8.5$No)
stationInfo = data.frame(stations=stations, lat=NA, long=NA)
unik <- which(!duplicated(dat8.5$No))
head(unik)
stationInfo$lat <- dat8.5$Lat[unik]
stationInfo$long <- dat8.5$Lon[unik]
names(stationInfo)[1] <- "No"
head(stationInfo)
which(!complete.cases(stationInfo))

#-------------------------------- 
### ICV matrix
head(ICV)  
C <- ICV %>% select(No, SST, Arag, pH)
head(C)
#--------------------------------   
```

```{r}
#---------------
### Hemisphere Indexes ###
# N. hemisphere
N_hem <- which(stationInfo$lat>=0)
S_hem <- which(stationInfo$lat<0)
#---------------
```


# 1800 analog to today (Novelty)

```{r}
#--------------------------------  
### 1800 analog to today (Novelty) ####
# What are today's novel climates compared to 1800?
# Novel climates are identified by comparing the 
# future climate realization (B) for each gridpoint 
# to the past (A) climate realizations (Williams)
# RESTRICT BASELINE TO HEMISPHERE OF OBSERVATION
#--------------------------------
# A is baseline, the base that a station from B is compared to
# If A is past and B is future, then degree of novelty
# If A is future and B is past, then degree of disappearance
# B will be subset to a particular station for analysis
# C is the data frame used to calculate the ICV, will also be subset to a particular station for analysis


A <- norm_1800
# 1800 climate normals
# Serves as the Baseline
head(A)
dim(A)

# subset the 2000 data to the same stations as the 1800 data
B <- norm_2000
# 2000 climate normals 
head(B)
dim(B)
# same columns as A

# sanity check to make sure stations in right order
identical(A$No, B$No) # should be true

NN.sigma_1800_today_N <- loop_sigma_D(A[N_hem,], B[N_hem,], C, "_1800_today")
NN.sigma_1800_today_S <- loop_sigma_D(A[S_hem,], B[S_hem,], C, "_1800_today")
NN.sigma_1800_today <- rbind(NN.sigma_1800_today_N, NN.sigma_1800_today_S)
head(NN.sigma_1800_today)
hist(NN.sigma_1800_today$numPCs_1800_today)

final_dat <- full_join(stationInfo, NN.sigma_1800_today,  by="No")
head(final_dat)

final_dat[which(is.infinite(final_dat$NN.sigma_1800_today)),]

final_dat[which(!complete.cases(final_dat)),]

identical(nrow(final_dat), nrow(stationInfo)) 
# should be true, same number as stations as stationInfo

# Visualize
world <- map_data("world2")
Plot_nonInt(final_dat$lat, final_dat$long, 
            final_dat$NN.sigma_1800_today, world, "sigD novel 2000")

# The NN.sigma_1800_today represents the novelty of today's
# climate compared to 1800
# NN.dist_1800_today represents the Mahalanobis distance from the climate
# in today to it's nearest neighbor in all 1800 data
# No--> NN.station_1800_today represents the station that was queried 2000 (B) 
# and it's nearest data in the global reference NN.station_1800_today (1800, A)
```

# Today analog to 1800 Disappearance 
```{r}
#--------------------------------  
### Today analog to 1800 Disappearance####
# What are 1800 disappearing climates?
# disappearing climates are identified by comparing
# each past gridpoint (A) to all future climate realizations (B)
#--------------------------------
calc_sigma_D(A, B, C, 1)
calc_sigma_D(B, A, C, 1)

B <- norm_1800
A <- norm_2000
identical(A$No, B$No) # should be true
identical(sort(A$No), A$No) # should be true


NN.sigma_today_1800_N <- loop_sigma_D(A[N_hem,], B[N_hem,], C, "_today_1800")
NN.sigma_today_1800_S <- loop_sigma_D(A[S_hem,], B[S_hem,], C, "_today_1800")
NN.sigma_today_1800 <- rbind(NN.sigma_today_1800_N,
                             NN.sigma_today_1800_S)

head(NN.sigma_today_1800)

final_dat2 <- full_join(final_dat, NN.sigma_today_1800)
head(final_dat2)
dim(final_dat2)
Plot_nonInt(final_dat2$lat, final_dat2$long, 
            final_dat2$NN.sigma_today_1800, world, "sigD dis 2000")

# The NN.sigma_1800_disappear represents the novelty of 1800
# compared to today (e.g disappearing climates)
# NN.dist_1800_disappear represents the Mahalanobis distance from the climate
# in 1800 to it's nearest neighbor in all of today's data
# No--> NN.station_1800_disappear represents the the station that was queried 1800 (B) 
# and it's nearest data in the global reference NN.station_1800_disappear (2000, A)
```

# Add the info for lat/long of nearest neighbors (1800 to today)
```{r}
stationInfo2 <- stationInfo
names(stationInfo2) <- paste0(names(stationInfo2),"_1800_today")
head(stationInfo2)
dim(stationInfo)
names(stationInfo2)[1] <-"NN.station_1800_today"
final_dat3 <- left_join(final_dat2, stationInfo2)
dim(final_dat3)
head(final_dat3)
tail(final_dat3)

# Should both be 0
sum(!complete.cases(final_dat2))
sum(!complete.cases(final_dat3))
head(final_dat3[!complete.cases(final_dat3),])
summary(final_dat3, NN.station_today_1800)


final_dat3$latshift_1800_today <- NA
condN <- which(final_dat3$lat>0)
final_dat3$latshift_1800_today[condN] <- final_dat3$lat[condN] - 
  final_dat3$lat_1800_today[condN]
condS <- which(final_dat3$lat<0)
final_dat3$latshift[condS] <- -1*(final_dat3$lat[condS] - 
                              final_dat3$lat_1800_today[condS])
hist(final_dat3$latshift_1800_today)

```

```{r}

write.csv(final_dat3, paste0(results_dir, "1800base_2000_novelty_NShemrestrict.csv"))
final_dat4 <- final_dat3
```


# Today analog to 2100 RCP 8.5 (Novelty) 
```{r}
identical(norm_2000$No, norm_2100_8.5$No)

A1 <- norm_2000
# 1970-2000 climate normals
head(A1)
dim(A1)

B1 <- norm_2100_8.5
# 2070-2100
head(B1)
dim(B1)
# same columns as A

# sanity check to make sure stations in right order
identical(A1$No, B1$No) # should be true
identical(A1$No, sort(A1$No)) # should also be true

# C is the same defined above
NN.sigma_today_2100_8.5_N <- loop_sigma_D(A1[N_hem,], B1[N_hem,], C, "_today_2100_8.5")
# C is the same defined above
NN.sigma_today_2100_8.5_S <- loop_sigma_D(A1[S_hem,], B1[S_hem,], C, "_today_2100_8.5")
NN.sigma_today_2100_8.5 <- rbind(NN.sigma_today_2100_8.5_N, 
                                 NN.sigma_today_2100_8.5_S)


which(is.infinite(NN.sigma_today_2100_8.5$NN.sigma_today_2100_8.5))
which(is.na(NN.sigma_today_2100_8.5$NN.sigma_today_2100_8.5))

final_dat5 <- full_join(NN.sigma_today_2100_8.5, final_dat4, by="No")
head(final_dat5)

dim(final_dat5)
stationInfo5 <- stationInfo
names(stationInfo5) <- paste0(names(stationInfo),"_today_2100_8.5")
names(stationInfo5)[1] <- "NN.station_today_2100_8.5"
head(stationInfo5)
final_dat6 <- left_join(final_dat5, stationInfo5)
dim(final_dat6)
head(final_dat6)

# Visualize
Plot_nonInt(final_dat6$lat, final_dat6$long, 
            final_dat6$NN.sigma_today_2100_8.5, world, "sigma nov 2100 8.5")
```

# 2100 RCP 8.5 to today, what are today's climates that  will disappear in 2100 RCP 8.5?
```{r}
#--------------------------------  
### 2100 RCP 8.5 to today, what are today's climates that ####
### will disappear in 2100 RCP 8.5?
#--------------------------------

NN.sigma_2100_8.5_today_N <- loop_sigma_D( B1[N_hem,], A1[N_hem,], C, "_2100_8.5_today")
NN.sigma_2100_8.5_today_S <- loop_sigma_D( B1[S_hem,], A1[S_hem,], C, "_2100_8.5_today")
NN.sigma_2100_8.5_today <- rbind(NN.sigma_2100_8.5_today_N, NN.sigma_2100_8.5_today_S)


final_dat7 <- full_join(NN.sigma_2100_8.5_today, final_dat6, by="No")
head(final_dat7)
Plot_nonInt(final_dat7$lat, final_dat7$long, 
            final_dat7$NN.sigma_2100_8.5_today, world, "sigma dis. 2000 in 2100 8.5")

final_dat8 <- final_dat7
```
# 2100 RCP 4.5 against today (novelty)
```{r}
#--------------------------------  
### 2100 RCP 4.5 against today (novelty) ####
#--------------------------------
identical(norm_2000$No, norm_2100_4.5$No)

# A1 the same as above

B2 <- norm_2100_4.5
# 2070-2100
head(B2)
dim(B2)
# same columns as A

# sanity check to make sure stations in right order
identical(A1$No, B2$No) # should be true
identical(A1$No, sort(A1$No)) # should also be true

# C is the same defined above
NN.sigma_today_2100_4.5_N <- loop_sigma_D(A1[N_hem,], B2[N_hem,], C, "_today_2100_4.5")
NN.sigma_today_2100_4.5_S <- loop_sigma_D(A1[S_hem,], B2[S_hem,], C, "_today_2100_4.5")
NN.sigma_today_2100_4.5 <- rbind(NN.sigma_today_2100_4.5_N, NN.sigma_today_2100_4.5_S)

which(is.infinite(NN.sigma_today_2100_4.5$NN.sigma_today_2100_4.5))
which(is.na(NN.sigma_today_2100_4.5$NN.sigma_today_2100_4.5))


final_dat9 <- full_join(NN.sigma_today_2100_4.5, final_dat8, by="No")
head(final_dat9)


# Visualize
Plot_nonInt(final_dat9$lat, final_dat9$long, 
            final_dat9$NN.sigma_today_2100_4.5, world, "sig nov 2100 4.5")


stationInfo9 <- stationInfo
names(stationInfo9) <- paste0(names(stationInfo),"_today_2100_4.5")
names(stationInfo9)[1] <- "NN.station_today_2100_4.5"
head(stationInfo9)
final_dat10 <- left_join(final_dat9, stationInfo9)
dim(final_dat10)
head(final_dat10)
```


# today against 2100 RCP 4.5 (disappearing climate)
```{r}
#--------------------------------  
### today against 2100 RCP 4.5 (disappearing climate) ####
#--------------------------------
NN.sigma_2100_4.5_today_N <- loop_sigma_D(B2[N_hem,], A1[N_hem,], C, "_2100_4.5_today")
NN.sigma_2100_4.5_today_S <- loop_sigma_D(B2[S_hem,], A1[S_hem,], C, "_2100_4.5_today")
NN.sigma_2100_4.5_today <- rbind(NN.sigma_2100_4.5_today_N, NN.sigma_2100_4.5_today_S)


which(is.infinite(NN.sigma_2100_4.5_today$NN.sigma_2100_4.5_today))
which(is.na(NN.sigma_2100_4.5_today$NN.sigma_2100_4.5_today))


final_dat11 <- full_join(NN.sigma_2100_4.5_today, final_dat10, by="No")
head(final_dat11)

# Visualize
Plot_nonInt(final_dat11$lat, final_dat11$long, 
            final_dat11$NN.sigma_2100_4.5_today, world, "sigma dis. 2000 in 2100 4.5")

final_dat12 <- final_dat11
dim(final_dat12)
head(final_dat12)

sum(!complete.cases(final_dat12))
```

```{r}
### Final dataframe ####
dat_Nov <- final_dat12
### Write to file ####
write.csv(final_dat12, paste0(results_dir, "SigmaD_Hem.csv"), row.names = FALSE)

head(dat_Nov)
for_liqing <- data.frame(No=dat_Nov$No, lat=dat_Nov$lat,long=dat_Nov$long,
                         A=dat_Nov$NN.sigma_today_1800, 
                         B=dat_Nov$NN.sigma_1800_today, 
                         C=dat_Nov$NN.sigma_2100_4.5_today, 
                         D=dat_Nov$NN.sigma_today_2100_4.5, 
                         E=dat_Nov$NN.sigma_2100_8.5_today, 
                         F=dat_Nov$NN.sigma_today_2100_8.5)
write.csv(for_liqing,  paste0(results_dir,"SigmaD_Hem_plot1.csv"), row.names=FALSE)

for_liqing_Md <- data.frame(No=dat_Nov$No, lat=dat_Nov$lat,long=dat_Nov$long,A=dat_Nov$NN.Mdist_today_1800, B=dat_Nov$NN.Mdist_1800_today, C=dat_Nov$NN.Mdist_2100_4.5_today, D=dat_Nov$NN.Mdist_today_2100_4.5, E=dat_Nov$NN.Mdist_2100_8.5_today, F=dat_Nov$NN.Mdist_today_2100_8.5)
write.csv(for_liqing_Md,  paste0(results_dir,"SigmaD_Hem_plot2.csv"), row.names=FALSE)
```

# Calculate percentages
```{r}
### Calculate percentages ####  
N <- nrow(dat_Nov)  

calc_perc <- function(x, descrip){
  print(paste("Percent of cells with moderate dissimilarity for", descrip,":", 
              round(sum(x>2 & x<4)/N,3)*100, sep=" "))
  print(paste("Percent of cells with extreme dissimilarity for", descrip,":", 
              round(sum(x>4)/N, 3)*100, sep=" "))
}

calc_perc(dat_Nov$NN.sigma_today_1800, "1800 disappearing in 2000")
calc_perc(dat_Nov$NN.sigma_1800_today, "Novel climates in 2000 since 1800")
calc_perc(dat_Nov$NN.sigma_2100_4.5_today, "2000 disappearing in 2100 4.5")
calc_perc(dat_Nov$NN.sigma_today_2100_4.5, "Novel climates in 2100 4.5 since 2000")
calc_perc(dat_Nov$NN.sigma_2100_8.5_today, "2000 disappearing in 2100 8.5")
calc_perc(dat_Nov$NN.sigma_today_2100_8.5, "Novel climates in 2100 8.5 since 2000")

```
