---
title: "NS Hemisphere visualize variability"
author: "KE Lotterhos"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To use this code, set your working directory to the /src folder

You will also have to update the paths below to be the paths to where the data is stored on your computer and where you want to output the results.

```{r}


#library(FNN)
#library(ggplot2)
#library(ggmap)
#library(mapdata)
#library(maps)

code_dir <- "/src"
setwd("code_dir")

results_dir <- ("../results/") # you may have to create

data_dir <- ("../Data/")

dat1 <- fread(paste0(data_dir,"model/Temp_Arag_1800_2000.txt"), sep = ",")

dat4.5 <- fread(paste0(data_dir,"model/Temp_Arag_2070_2100_RCP45.txt"), sep = ",")
                
dat8.5 <- fread(paste0(data_dir,"model/Temp_Arag_2070_2100_RCP85.txt"), sep = ",")

dat_2000 <-   dat1 %>% filter(Year>1960 & Year<2010)
dat_1800 <- dat1 %>% filter(Year<1850)

setwd(code_dir)
source("0-Novelty_Oceans_Functions.R")
```

## Set colors
```{r}
c1800 = "black"
c2000 = "darkblue"
c4.5 = "orange"
c8.5 = "brown1"

```


## Combine climate envelope legend
```{r, Combine climate envelope legend}
## Create a 4-color legend
legend_data <- data.frame(Year=c("1800","2000",
                                 "2100 RCP 4.5",
                                 "2100 RCP 8.5"),
                          Color = adjustcolor(c(c1800,c2000, c4.5, c8.5), alpha=0.5),
                          val=c(1,2,3,4)
)



g_leg <- ggplot(legend_data, aes(x=Year,y=val, fill=Year)) + geom_hex() + scale_fill_manual(name="Year",
                  values=adjustcolor(c(c1800,c2000, c4.5, c8.5), alpha=0.5)) + theme(legend.position = "right", legend.text = element_text(size=20), legend.key.size = unit(0.5, "in"), legend.title = element_text(size=20))


get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(g_leg)
```

## Lat vs. SST
```{r}

all <- ggplot() + 
  theme_classic() + 
theme(axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), axis.title = element_text(size=18), plot.title=element_text(size=16)) 
#+  scale_y_continuous(
#  labels = scales::number_format(accuracy = 0.1))

g0 <- all + 
  geom_hex(data= dat_1800, aes(Lat, SST), fill="black", alpha=0.5)  + ylim(-5,35) +
  geom_hex(data= dat_2000, aes(Lat, SST), fill="darkblue", alpha=0.5) +
  ggtitle("A) 1800 vs. 2000 SST") +  theme(axis.title.x = element_blank())

g1 <- all +
  geom_hex(data= dat_2000, aes(Lat, SST), fill="darkblue", alpha=0.5) + ylim(-5,35) + theme(axis.title.x = element_blank())
#g1

g2 <- g1 +  geom_hex(data= dat4.5, aes(Lat, SST), alpha=0.5, fill="orange") +
  ggtitle("B)  2000 vs. 2100 RCP 4.5 SST")
#g2

g3 <- g1 +  geom_hex(data= dat8.5, aes(Lat, SST), alpha=0.5, fill="brown1") +
  ggtitle("C)  2000 vs. 2100 RCP 8.5 SST")
#g3
```



## Lat vs. Arag Sat State
```{r}


a0 <-  all +
  geom_hex(data= dat_1800, aes(Lat, Arag), fill="black", alpha=0.5)  + 
  geom_hex(data= dat_2000, aes(Lat, Arag), fill="darkblue", alpha=0.5) +
  ggtitle("D) 1800 vs 2000 Arag") +  scale_y_continuous(limits = c(0.2,5.5), labels = scales::number_format(accuracy = 0.1)) + theme(axis.title.x = element_blank())

#a0

a1 <- all +
  geom_hex(data= dat_2000, aes(Lat, Arag), fill="darkblue", alpha=0.5) +  scale_y_continuous(limits = c(0.2,5.5), labels = scales::number_format(accuracy = 0.1)) + theme(axis.title.x = element_blank())
#a1

a2 <- a1 +  geom_hex(data= dat4.5, aes(Lat, Arag), alpha=0.5, fill="orange") +
  ggtitle("E) 2000 vs. 2100 RCP 4.5 Arag")
#a2

a3 <- a1 +  geom_hex(data= dat8.5, aes(Lat, Arag), alpha=0.5, fill="brown1") +
  ggtitle("F) 2000 vs. 2100 RCP 8.5 Arag")
#a3
```

## Lat vs. pH
```{r}
p0 <- all + 
  geom_hex(data= dat_1800, aes(Lat, pH), fill="black", alpha=0.5) + 
  geom_hex(data= dat_2000, aes(Lat, pH), fill="darkblue", alpha=0.5) +
  ggtitle("G) 1800 vs 2000 pH") + scale_y_continuous(limits = c(7.2,8.75),
  labels = scales::number_format(accuracy = 0.1))
p0

p1 <- all + 
  geom_hex(data= dat_2000, aes(Lat, pH), fill="darkblue", alpha=0.5)  + scale_y_continuous(limits = c(7.2,8.75),
  labels = scales::number_format(accuracy = 0.1))
#p1

p2 <- p1 +  geom_hex(data= dat4.5, aes(Lat, pH), alpha=0.5, fill="orange") +
  ggtitle("H) 2000 vs. 2100 RCP 4.5 pH")
#p2

p3 <- p1 +  geom_hex(data= dat8.5, aes(Lat, pH), alpha=0.5, fill="brown1") +
  ggtitle("I) 2000 vs. 2100 RCP 8.5 pH")
#p3
```


```{r}
pdf("../figures_hem/Lat_figs.pdf",
    width=15, height=12)
grid.arrange(g0, g2, g3, 
             a0, a2, a3,
             p0, p2, p3, legend,
             nrow=3, ncol=4,
             layout_matrix = rbind(c(1,2,3, 10), c(4,5,6, 10), c(7,8,9,10)),
             widths = c(rep(12/3,3), 3), heights = rep(12/3,3))
dev.off()
```


## Multiple variable plots

* upper left: Temp vs Arag for N hemisphere, 1800 vs 2000
* upper right: Temp vs Arag for N hemisphere, 2000 vs 2100
* lower left: Temp vs pH for N hemisphere, 1800 vs 2000
* lower right: Temp vs pH for N hemisphere, 2000 vs 2100

(Repeat for S hemisphere)

### Set up for S and N hemisphere envelopes

This code needs to be run twice, once for the N hem and once for the S hem

```{r, N hemisphere envelopes Arag}

N_hem_1800 <- (dat_1800$Lat>=0)
N_hem_2000 <- (dat_2000$Lat>=0)
N_hem_4.5 <- (dat4.5$Lat>=0)
N_hem_8.5 <- (dat8.5$Lat>=0)


run_North = TRUE


if(run_North==TRUE){
  plotind_1800 = which(N_hem_1800)
  plotind_2000 = which(N_hem_2000)
  plotind_4.5 = which(N_hem_4.5)
  plotind_8.5 = which(N_hem_8.5)
}


all2 <- ggplot() + theme_classic() +
theme(axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), axis.title = element_text(size=18), plot.title = element_text(size=16)) 

```

Aragonite vs SST N hemisphere

```{r}


e0N <-  all2 + geom_hex(data= dat_2000[plotind_2000,], aes(SST, Arag), fill=c2000, alpha=0.5) + labs(y="Aragonite Saturation State") + geom_hline(aes(yintercept=1), linetype =2) + scale_y_continuous(limits=c(0.2, 5.5),  labels = scales::number_format(accuracy = 0.1))
#e0N

e1N <- e0N +  geom_hex(data= dat_1800[plotind_1800,], aes(SST, Arag), fill=c1800, alpha=0.5) +
  ggtitle("A) 1800 vs 2000, N. Hem") + 
  annotate(geom="text", x=0, y=5, label="A) 1800 disappearing in 2000", adj=0, color="grey30") +
  annotate(geom="text", x=20, y=1.3, label="Novel in 2000", adj=0, color=adjustcolor(c2000,0.5)) + 
  geom_segment(aes(x=7, y = 4.8, xend=0, yend=4), color="grey30") + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=4), color="grey30") +
  geom_segment(aes(x=20, y = 1.4, xend=11, yend=1.7), color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=20, y = 1.4, xend=23, yend=2.8), color=adjustcolor(c2000,0.5)) 
#e1N

e2N <- e0N +  geom_hex(data= dat4.5[plotind_4.5,], aes(SST, Arag), fill=c4.5, alpha=0.5) +
  ggtitle("B) 2000 vs 2100 RCP 4.5, N. Hem") + 
  annotate(geom="text", x=0, y=5, label="2000 disappearing in 2100 RCP 4.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=0, yend=2.5), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=3.5), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text", x=10, y=0.5, label="Novel in 2100 RCP 4.5", adj=0, color=adjustcolor(c4.5,0.9))  +
  geom_segment(aes(x=15, y = 0.6, xend=11, yend=1.5), color=adjustcolor(c4.5,0.9)) +
  geom_segment(aes(x=15, y = 0.6, xend=23, yend=2.5), color=adjustcolor(c4.5,0.9)) 
#e2N  

e3N <- e0N +  geom_hex(data= dat8.5[plotind_8.5,], aes(SST, Arag), fill=c8.5, alpha=0.5) +
  ggtitle("C) 2000 vs 2100 RCP 8.5, N. Hem") + 
  annotate(geom="text", x=0, y=5, label="2000 disappearing in 2100 RCP 8.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=0, yend=2.2), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=3.0), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text", x=12, y=0.2, label="Novel in 2100 RCP 8.5", adj=0, color=adjustcolor(c8.5,0.9))  +
  geom_segment(aes(x=20, y = 0.3, xend=11, yend=1.3), color=adjustcolor(c8.5,0.9)) +
  geom_segment(aes(x=20, y = 0.3, xend=23, yend=2.1), color=adjustcolor(c8.5,0.9)) 
#e3N  
```

pH vs SST N hemisphere

```{r, N hemisphere envelopes pH}


p0N <- all2+ ylim(7.2,8.75) + geom_hex(data= dat_2000[plotind_2000,], aes(SST, pH), fill=c2000, alpha=0.5)
#p0N

p1N <- p0N +  geom_hex(data= dat_1800[plotind_1800,], aes(SST, pH), fill=c1800, alpha=0.5) +
  ggtitle("D) 1800 vs 2000, N. Hem") + 
  annotate(geom="text", x=7, y=8.7, label="1800 disappearing in 2000", adj=0, color="grey30") + 
  geom_segment(aes(x=15, y = 8.65, xend=8, yend=8.4), color="grey30") + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.25), color="grey30") +
  annotate(geom="text", x=15, y=7.6, label="Novel in 2000", adj=0, color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=15, y = 7.63, xend=3, yend=7.8), color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=15, y = 7.63, xend=21, yend=8.0), color=adjustcolor(c2000,0.5)) 
#p1N

p2N <- p0N +  geom_hex(data= dat4.5[plotind_4.5,], aes(SST, pH), fill=c4.5, alpha=0.5) +
  ggtitle("E) 2000 vs 2100 RCP 4.5, N. Hem") + 
  annotate(geom="text",  x=5, y=8.7, label="2000 disappearing in 2100 RCP 4.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=8, yend=8.2), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.1), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text",  x=15, y=7.6, label="Novel in 2100 RCP 4.5", adj=0, color=adjustcolor(c4.5,0.9))  +
  geom_segment(aes(x=15, y = 7.63, xend=3, yend=7.6), color=adjustcolor(c4.5,0.9)) +
  geom_segment(aes(x=15, y = 7.63, xend=21, yend=7.9), color=adjustcolor(c4.5,0.9)) 
#p2N  

p3N <- p0N +  geom_hex(data= dat8.5[plotind_8.5,], aes(SST, pH), fill=c8.5, alpha=0.5) +
  ggtitle("F) 2000 vs 2100 RCP 8.5, N. Hem") + 
  annotate(geom="text",  x=5, y=8.7, label="2000 disappearing in 2100 RCP 8.5", adj=0, color=adjustcolor(c2000,0.7)) +
  geom_segment(aes(x=15, y = 8.65, xend=8, yend=8.2), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.1), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text",  x=15, y=7.5, label="Novel in 2100 RCP 8.5", adj=0, color=adjustcolor(c8.5,0.9))  +
  geom_segment(aes(x=15, y = 7.53, xend=3, yend=7.6), color=adjustcolor(c8.5,0.9)) +
  geom_segment(aes(x=15, y = 7.53, xend=21, yend=7.8), color=adjustcolor(c8.5,0.9)) 
#p3N
```


```{r}
if(run_North==TRUE){filename="../figures_hem/envelopes_N_hem.pdf"}

pdf(filename,
    width=16, height=8)
grid.arrange(e1N, e2N, e3N, 
             p1N, p2N, p3N, legend,
             nrow=2, ncol=4,
             layout_matrix = rbind(c(1,2,3, 7), c(4,5,6, 7)),
             widths = c(rep(13/3,3), 3), heights = c(4,4))
dev.off()
```

## S hemisphere envelopes
```{r, S hem envelopes}
run_North = FALSE
if(run_North==FALSE){
  plotind_1800 = which(!N_hem_1800)
  plotind_2000 = which(!N_hem_2000)
  plotind_4.5 = which(!N_hem_4.5)
  plotind_8.5 = which(!N_hem_8.5)
}
if(run_North==FALSE){filename="../figures_hem/envelopes_S_hem.pdf"}

e0N <-  all2 + geom_hex(data= dat_2000[plotind_2000,], aes(SST, Arag), fill=c2000, alpha=0.5) + labs(y="Aragonite Saturation State") + geom_hline(aes(yintercept=1), linetype =2) + scale_y_continuous(limits=c(0.2, 5.5),  labels = scales::number_format(accuracy = 0.1))
#e0N

e1N <- e0N +  geom_hex(data= dat_1800[plotind_1800,], aes(SST, Arag), fill=c1800, alpha=0.5) +
  ggtitle("A) 1800 vs 2000, S. Hem") + 
  annotate(geom="text", x=0, y=5, label="A) 1800 disappearing in 2000", adj=0, color="grey30") +
  annotate(geom="text", x=20, y=1.3, label="Novel in 2000", adj=0, color=adjustcolor(c2000,0.5)) + 
  geom_segment(aes(x=7, y = 4.8, xend=-2, yend=4), color="grey30") + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=4), color="grey30") +
  geom_segment(aes(x=20, y = 1.4, xend=11, yend=1.6), color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=20, y = 1.4, xend=23, yend=2.8), color=adjustcolor(c2000,0.5)) 
#e1N

e2N <- e0N +  geom_hex(data= dat4.5[plotind_4.5,], aes(SST, Arag), fill=c4.5, alpha=0.5) +
  ggtitle("B) 2000 vs 2100 RCP 4.5, S. Hem") + 
  annotate(geom="text", x=0, y=5, label="2000 disappearing in 2100 RCP 4.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=-2, yend=2.5), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=3.5), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text", x=10, y=0.5, label="Novel in 2100 RCP 4.5", adj=0, color=adjustcolor(c4.5,0.9))  +
  geom_segment(aes(x=15, y = 0.6, xend=11, yend=1.15), color=adjustcolor(c4.5,0.9)) +
  geom_segment(aes(x=15, y = 0.6, xend=23, yend=2.2), color=adjustcolor(c4.5,0.9)) 
#e2N  

e3N <- e0N +  geom_hex(data= dat8.5[plotind_8.5,], aes(SST, Arag), fill=c8.5, alpha=0.5) +
  ggtitle("C) 2000 vs 2100 RCP 8.5, S. Hem") + 
  annotate(geom="text", x=0, y=5, label="2000 disappearing in 2100 RCP 8.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=0, yend=2.2), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=7, y = 4.8, xend=20, yend=3.0), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text", x=12, y=0.2, label="Novel in 2100 RCP 8.5", adj=0, color=adjustcolor(c8.5,0.9))  +
  geom_segment(aes(x=20, y = 0.3, xend=11, yend=1.0), color=adjustcolor(c8.5,0.9)) +
  geom_segment(aes(x=20, y = 0.3, xend=23, yend=2.1), color=adjustcolor(c8.5,0.9)) 
#e3N  

## ph ####


p0N <- all2+ ylim(7.2,8.75) + geom_hex(data= dat_2000[plotind_2000,], aes(SST, pH), fill=c2000, alpha=0.5)
#p0N

p1N <- p0N +  geom_hex(data= dat_1800[plotind_1800,], aes(SST, pH), fill=c1800, alpha=0.5) +
  ggtitle("D) 1800 vs 2000, S. Hem") + 
  annotate(geom="text", x=7, y=8.7, label="1800 disappearing in 2000", adj=0, color="grey30") + 
  geom_segment(aes(x=15, y = 8.65, xend=5, yend=8.25), color="grey30") + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.25), color="grey30") +
  annotate(geom="text", x=15, y=7.6, label="Novel in 2000", adj=0, color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=15, y = 7.63, xend=7, yend=8.05), color=adjustcolor(c2000,0.5)) +
  geom_segment(aes(x=15, y = 7.63, xend=25, yend=8.0), color=adjustcolor(c2000,0.5)) 
#p1N

p2N <- p0N +  geom_hex(data= dat4.5[plotind_4.5,], aes(SST, pH), fill=c4.5, alpha=0.5) +
  ggtitle("E) 2000 vs 2100 RCP 4.5, S. Hem") + 
  annotate(geom="text",  x=5, y=8.7, label="2000 disappearing in 2100 RCP 4.5", adj=0, color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=8, yend=8.1), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.1), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text",  x=15, y=7.6, label="Novel in 2100 RCP 4.5", adj=0, color=adjustcolor(c4.5,0.9))  +
  geom_segment(aes(x=15, y = 7.63, xend=3, yend=7.9), color=adjustcolor(c4.5,0.9)) +
  geom_segment(aes(x=15, y = 7.63, xend=21, yend=7.85), color=adjustcolor(c4.5,0.9)) 
#p2N  

p3N <- p0N +  geom_hex(data= dat8.5[plotind_8.5,], aes(SST, pH), fill=c8.5, alpha=0.5) +
  ggtitle("F) 2000 vs 2100 RCP 8.5, S. Hem") + 
  annotate(geom="text",  x=5, y=8.7, label="2000 disappearing in 2100 RCP 8.5", adj=0, color=adjustcolor(c2000,0.7)) +
  geom_segment(aes(x=15, y = 8.65, xend=8, yend=8.1), color=adjustcolor(c2000,0.7)) + 
  geom_segment(aes(x=15, y = 8.65, xend=23, yend=8.1), color=adjustcolor(c2000,0.7)) +
  annotate(geom="text",  x=15, y=7.5, label="Novel in 2100 RCP 8.5", adj=0, color=adjustcolor(c8.5,0.9))  +
  geom_segment(aes(x=15, y = 7.53, xend=3, yend=7.8), color=adjustcolor(c8.5,0.9)) +
  geom_segment(aes(x=15, y = 7.53, xend=21, yend=7.8), color=adjustcolor(c8.5,0.9)) 

### plot ####

pdf(filename,
    width=16, height=8)
grid.arrange(e1N, e2N, e3N, 
             p1N, p2N, p3N, legend,
             nrow=2, ncol=4,
             layout_matrix = rbind(c(1,2,3, 7), c(4,5,6, 7)),
             widths = c(rep(13/3,3), 3), heights = c(4,4))
dev.off()
```